---
title: "Geocoding Evaluation: Honest Held-Out Accuracy & Calibration"
subtitle: "Leakage-controlled evaluation harness (spec 2026-07, issue #47)"
author: "Pipeline Automated Report"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-fold: true
    theme: cosmo
    fig-width: 10
    fig-height: 6
    self-contained: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(data.table)
library(ggplot2)
library(gt)
library(targets)

theme_set(theme_minimal(base_size = 12))

tar_load(c(
  tse_coverage,
  accuracy_tables,
  calibration_check
))

acc <- as.data.table(accuracy_tables)
cov <- as.data.table(tse_coverage)
```

## What this report measures {.unnumbered}

::::: {.callout-important}
**The stations the model actually determines are not measured here.** Wherever the
TSE has a coordinate, the pipeline's output *is* that TSE coordinate; the
model-selected coordinate only reaches the output for TSE-**uncovered** stations.
So every accuracy number below is measured on the TSE-**covered** subset and is an
**extrapolation** to the uncovered subset. The uncovered subset is reported as
explicitly unmeasured, not folded silently into any headline.

Accuracy is measured **out-of-fold**: each covered station is scored by a model
refit on other municipalities' folds, never on its own (spec section 3). Error is
the haversine distance (km) from the model-selected coordinate to the
field-collected TSE coordinate. Accuracy and match rate are always reported
jointly (they trade off). Strata below the held-out N floor are suppressed.
:::::

```{r}
#| label: headline
#| output: asis
overall <- acc[stratum == "overall"]
cat(sprintf(
  paste0(
    "**Headline (covered set):** median error **%.3f km**, ",
    "**%.1f%% within 500 m**, match rate **%.1f%%** ",
    "over **%s** covered stations (**%s** geocoded)."
  ),
  overall$median_km, overall$within_500m, overall$match_rate,
  format(overall$n_total, big.mark = ","),
  format(overall$n_geocoded, big.mark = ",")
))
```

## Accuracy by stratum

```{r}
#| label: acc-table
acc_display <- copy(acc)
num_cols <- c("match_rate", "median_km", "p90", "p95", "p99",
              "within_100m", "within_500m", "within_1km")
acc_display |>
  gt(groupname_col = "stratum", rowname_col = "level") |>
  fmt_number(columns = c("median_km", "p90", "p95", "p99"), decimals = 3) |>
  fmt_number(columns = c("match_rate", "within_100m", "within_500m", "within_1km"),
             decimals = 1) |>
  fmt_integer(columns = c("n_total", "n_geocoded")) |>
  sub_missing(columns = num_cols, missing_text = "—") |>
  cols_label(
    n_total = "N covered", n_geocoded = "N geocoded", match_rate = "Match %",
    median_km = "Median km", within_100m = "<100m %", within_500m = "<500m %",
    within_1km = "<1km %"
  ) |>
  tab_header(
    title = "Positional accuracy on the TSE-covered set (out-of-fold)",
    subtitle = "Suppressed strata (below held-out N floor) show counts only."
  ) |>
  tab_source_note("Error in kilometres. Median is the 50th percentile; p90/p95/p99 shown.")
```

## TSE coverage density (by year × state)

Each accuracy number should be read against how much ground truth its stratum has.
TSE coordinates begin with the 2018 vintage; pre-2018 station-years have no TSE
reference and are unmeasurable.

```{r}
#| label: coverage
cov_wide <- dcast(cov, sg_uf ~ ano, value.var = "coverage_pct")
cov_wide |>
  gt(rowname_col = "sg_uf") |>
  fmt_number(columns = where(is.numeric), decimals = 1) |>
  sub_missing(missing_text = "—") |>
  tab_header(title = "TSE coverage (% of covered stations)") |>
  tab_source_note("Share of polling stations with a non-missing TSE coordinate.")
```

## `pred_dist` calibration

### Rank-and-filter

Sorting covered stations by predicted error and dropping the worst-predicted tail
should monotonically lower realized median error and raise %-within-500 m. This
demonstrates the ranking carries information even where it is not calibrated in
absolute metres.

```{r}
#| label: rank-filter
rf <- as.data.table(calibration_check$rank_filter)
rf |>
  gt() |>
  fmt_number(columns = c("retained_frac"), decimals = 2) |>
  fmt_number(columns = c("median_km"), decimals = 3) |>
  fmt_number(columns = c("within_500m"), decimals = 1) |>
  fmt_integer(columns = "n_retained") |>
  cols_label(
    drop_frac = "Worst-predicted dropped", retained_frac = "Retained frac",
    n_retained = "N retained", median_km = "Median km", within_500m = "<500m %"
  ) |>
  tab_header(title = "Rank-and-filter on predicted error")

ggplot(rf, aes(x = drop_frac)) +
  geom_line(aes(y = median_km), color = "#007bff") +
  geom_point(aes(y = median_km), color = "#007bff") +
  labs(x = "Fraction of worst-predicted stations dropped",
       y = "Realized median error (km)",
       title = "Dropping the worst-predicted tail lowers realized error")
```

### Reliability & ENCE

```{r}
#| label: reliability
rel <- as.data.table(calibration_check$reliability)
ence <- calibration_check$ence
```

Expected Normalized Calibration Error (ENCE): **`r sprintf("%.3f", ence)`**
(0 = perfectly calibrated point estimate).

```{r}
#| label: reliability-plot
lim <- range(c(rel$mean_pred, rel$mean_realized), na.rm = TRUE)
ggplot(rel, aes(x = mean_pred, y = mean_realized)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(aes(size = n), color = "#007bff") +
  coord_equal(xlim = lim, ylim = lim) +
  labs(x = "Mean predicted error (km)", y = "Mean realized error (km)",
       size = "N", title = "Reliability: predicted vs realized error by bin")
```

## Google reference-validation (spec section 8)

Quantifies Google's own error budget relative to field-GPS TSE, on a covered-only
sample, *before* trusting Google for the uncovered subset. This is a
reference-validation, **not** a release gate and **not** a pipeline accuracy
number. The Google results are a frozen artifact (`data/google_geocoded.csv`);
regenerate it once with `scripts/generate_google_reference.R` (needs a Google
Geocoding API key).

```{r}
#| label: google
#| output: asis
google_file <- "../data/google_geocoded.csv"
if (!file.exists(google_file)) google_file <- "data/google_geocoded.csv"

if (file.exists(google_file)) {
  g <- fread(google_file)
  cat(sprintf(
    "Frozen Google artifact found: **%s** points. See table below.\n\n",
    format(nrow(g), big.mark = ",")
  ))
} else {
  cat(paste0(
    "_No frozen Google artifact found. This section is skipped until ",
    "`scripts/generate_google_reference.R` is run once to produce ",
    "`data/google_geocoded.csv`._\n"
  ))
}
```

```{r}
#| label: google-table
if (file.exists(google_file) && "google_long" %in% names(g) &&
    all(c("tse_long", "tse_lat") %in% names(g))) {
  g[, google_dist := geosphere::distHaversine(
    cbind(tse_long, tse_lat), cbind(google_long, google_lat)
  ) / 1000]
  by_col <- if ("urban_rural" %in% names(g)) "urban_rural" else NULL
  g_summary <- g[, .(
    n = .N,
    median_km = median(google_dist, na.rm = TRUE),
    within_500m = 100 * mean(google_dist <= 0.5, na.rm = TRUE)
  ), by = by_col]
  gt(g_summary) |>
    fmt_number(columns = "median_km", decimals = 3) |>
    fmt_number(columns = "within_500m", decimals = 1) |>
    tab_header(title = "Google-vs-TSE agreement (reference validation)")
}
```
