---
title: "Geocoding Evaluation: Honest Held-Out Accuracy & Calibration"
subtitle: "Leakage-controlled evaluation harness (spec 2026-07, issue #47)"
author: "Pipeline Automated Report"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-fold: true
    theme: cosmo
    fig-width: 10
    fig-height: 6
    self-contained: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(data.table)
library(ggplot2)
library(gt)
library(targets)

theme_set(theme_minimal(base_size = 12))

tar_load(c(
  tse_coverage,
  accuracy_tables,
  baseline_accuracy_tables,
  baseline_comparison,
  geocodebr_accuracy_tables,
  geocodebr_vs_model,
  calibration_check
))

acc <- as.data.table(accuracy_tables)
cov <- as.data.table(tse_coverage)
cmp <- as.data.table(baseline_comparison)
gb_acc <- as.data.table(geocodebr_accuracy_tables)
gb_cmp <- as.data.table(geocodebr_vs_model)

# One formatter per table schema, so every rendering of a schema agrees on decimals,
# labels and spanners. Each caller adds only its own filter, counts and notes.
acc_ladder_gt <- function(dt, title, subtitle) {
  dt |>
    gt(groupname_col = "stratum", rowname_col = "level") |>
    fmt_number(columns = c("median_km", "p90", "p95", "p99"), decimals = 3) |>
    fmt_number(columns = c("match_rate", "within_100m", "within_500m", "within_1km"),
               decimals = 1) |>
    fmt_integer(columns = c("n_total", "n_geocoded")) |>
    sub_missing(missing_text = "—") |>
    cols_label(
      n_total = "N covered", n_geocoded = "N geocoded", match_rate = "Match %",
      median_km = "Median km", within_100m = "<100m %", within_500m = "<500m %",
      within_1km = "<1km %"
    ) |>
    tab_header(title = title, subtitle = subtitle)
}

head_to_head_gt <- function(dt, title, subtitle = NULL) {
  dt |>
    gt(groupname_col = "universe", rowname_col = "level") |>
    fmt_number(
      columns = c("median_km_geocodebr", "median_km_model", "delta_median_km"),
      decimals = 3
    ) |>
    fmt_number(
      columns = c("within_500m_geocodebr", "within_500m_model", "delta_within_500m"),
      decimals = 1
    ) |>
    fmt_integer(columns = where(is.integer)) |>
    sub_missing(missing_text = "—") |>
    cols_label(
      median_km_geocodebr = "geocodebr", median_km_model = "Pipeline",
      delta_median_km = "Δ",
      within_500m_geocodebr = "geocodebr", within_500m_model = "Pipeline",
      delta_within_500m = "Δ"
    ) |>
    tab_spanner(
      label = "Median error (km)",
      columns = c("median_km_geocodebr", "median_km_model", "delta_median_km")
    ) |>
    tab_spanner(
      label = "Within 500 m (%)",
      columns = c("within_500m_geocodebr", "within_500m_model", "delta_within_500m")
    ) |>
    tab_header(title = title, subtitle = subtitle)
}
```

## What this report measures {.unnumbered}

::::: {.callout-important}
**The stations the model actually determines are not measured here.** Wherever the
TSE has a coordinate, the pipeline's output *is* that TSE coordinate; the
model-selected coordinate only reaches the output for TSE-**uncovered** stations.
So every accuracy number below is measured on the TSE-**covered** subset and is an
**extrapolation** to the uncovered subset. The uncovered subset is reported as
explicitly unmeasured, not folded silently into any headline.

Accuracy is measured **out-of-fold**: each covered station is scored by a model
refit on other municipalities' folds, never on its own (spec section 3). Error is
the haversine distance (km) from the model-selected coordinate to the
field-collected TSE coordinate. Accuracy and match rate are always reported
jointly (they trade off). Strata below the held-out N floor are suppressed.

*One caveat on "leakage-controlled": the per-fold refits are leak-free, but the
model hyperparameters were tuned once on the production split and reused across
folds rather than re-tuned per fold. This leaves a weak residual
hyperparameter-selection channel — a deliberately accepted simplification over
nested per-fold tuning, so these numbers may be marginally optimistic relative to
a fully nested estimate.*
:::::

```{r}
#| label: headline
#| output: asis
overall <- acc[stratum == "overall"]
cat(sprintf(
  paste0(
    "**Headline (covered set):** median error **%.3f km**, ",
    "**%.1f%% within 500 m**, match rate **%.1f%%** ",
    "over **%s** covered stations (**%s** geocoded)."
  ),
  overall$median_km, overall$within_500m, overall$match_rate,
  format(overall$n_total, big.mark = ","),
  format(overall$n_geocoded, big.mark = ",")
))
```

## Accuracy by stratum

```{r}
#| label: acc-table
acc |>
  acc_ladder_gt(
    title = "Positional accuracy on the TSE-covered set (out-of-fold)",
    subtitle = "Suppressed strata (below held-out N floor) show counts only."
  ) |>
  tab_source_note("Error in kilometres. Median is the 50th percentile; p90/p95/p99 shown.")
```

## Does the selection model beat a trivial heuristic?

The pipeline picks each station's coordinate with a tuned LightGBM model. This section
asks whether that complexity pays rent, by scoring a **trivial deterministic rule** on
the same candidates, the same covered stations, and the same protocol: take the
highest-precedence candidate available for the station — INEP school (by name), CNEFE
school, INEP school (by address), geocodebr, street median, neighborhood centroid —
breaking ties within a rank on the smallest string distance. Census vintages of the same
reference share a rank, so the string distance picks between them. The heuristic trains
on nothing, so it has nothing to hold out.

Both selectors rank the same candidate rows, so a station geocodes under one exactly
when it geocodes under the other: match rates are identical by construction and the
comparison is pure accuracy. Negative `Δ median` and positive `Δ <500 m` mean the model
is better.

```{r}
#| label: baseline-headline
#| output: asis
cmp_overall <- cmp[stratum == "overall"]
cat(sprintf(
  paste0(
    "**Overall (covered set):** median error **%.3f km** (model) vs **%.3f km** ",
    "(heuristic), a difference of **%+.3f km**. Within 500 m: **%.1f%%** vs ",
    "**%.1f%%**, a difference of **%+.1f pp**."
  ),
  cmp_overall$median_km_model, cmp_overall$median_km_baseline,
  cmp_overall$delta_median_km,
  cmp_overall$within_500m_model, cmp_overall$within_500m_baseline,
  cmp_overall$delta_within_500m
))
```

```{r}
#| label: baseline-table
cmp |>
  gt(groupname_col = "stratum", rowname_col = "level") |>
  fmt_number(
    columns = c("median_km_baseline", "median_km_model", "delta_median_km"),
    decimals = 3
  ) |>
  fmt_number(
    columns = c("within_500m_baseline", "within_500m_model", "delta_within_500m"),
    decimals = 1
  ) |>
  fmt_integer(columns = c("n_total", "n_geocoded")) |>
  sub_missing(missing_text = "—") |>
  cols_label(
    n_total = "N covered", n_geocoded = "N geocoded",
    median_km_baseline = "Heuristic", median_km_model = "Model",
    delta_median_km = "Δ",
    within_500m_baseline = "Heuristic", within_500m_model = "Model",
    delta_within_500m = "Δ", suppressed = "Below N floor"
  ) |>
  tab_spanner(
    label = "Median error (km)",
    columns = c("median_km_baseline", "median_km_model", "delta_median_km")
  ) |>
  tab_spanner(
    label = "Within 500 m (%)",
    columns = c("within_500m_baseline", "within_500m_model", "delta_within_500m")
  ) |>
  tab_header(
    title = "Selection model vs trivial source-precedence heuristic",
    subtitle = "Out-of-fold model picks; both scored on the same covered stations."
  ) |>
  tab_source_note(
    "Δ is model minus heuristic: negative median and positive <500 m favour the model."
  )
```

```{r}
#| label: baseline-source-mix
bl_src <- as.data.table(baseline_accuracy_tables)[stratum == "match_source",
  .(level, n_geocoded)]
model_src <- acc[stratum == "match_source", .(level, n_geocoded)]
src <- merge(model_src, bl_src, by = "level", all = TRUE,
             suffixes = c("_model", "_heuristic"))
src |>
  gt(rowname_col = "level") |>
  fmt_integer(columns = where(is.numeric)) |>
  sub_missing(missing_text = "—") |>
  cols_label(n_geocoded_model = "Model", n_geocoded_heuristic = "Heuristic") |>
  tab_header(
    title = "Which source each selector picks",
    subtitle = "Stations assigned to each candidate type. Same stations, different picks."
  )
```

## Could `geocodebr` replace the bespoke 2022 CNEFE tables?

`geocodebr` distributes IBGE's own aggregation of the 2022 CNEFE — the same source the
project's hand-built 2022 street and neighborhood reference tables are made from. If its
coordinates are as close to TSE ground truth as those tables' are, the tables can be
retired. This section runs that comparison. Nothing else is at stake: the 2010/2017
multi-vintage references, INEP school matching, the selection model, panel linkage and
station-specific normalization are all outside `geocodebr`'s scope regardless.

Unlike the heuristic above, `geocodebr` and the pipeline do **not** geocode the same
stations — each resolves some the other cannot. So every cell below is measured on the
stations where **both** produced a coordinate (`N both`), with each side's coverage of the
full cell shown alongside: parity on the intersection is worth nothing if `geocodebr`
resolves far fewer stations. **Δ is `geocodebr` minus the model** — the reverse of the
heuristic table above, since here the pipeline is the incumbent — so a negative median Δ
and a positive `<500 m` Δ favour **`geocodebr`**.

One caveat on reading `N geocodebr`: the cascade bottoms out at the municipal centroid, so
`geocodebr` returns *something* for nearly every station it is handed and its match rate is
close to 100% by construction. That is not a coverage win — a municipal centroid is a
coordinate, not a located station. The precision-tier table below is where the real coverage
question lives.

```{r}
#| label: geocodebr-headline
#| output: asis
gate <- gb_cmp[universe == "cnefe22_winner" & stratum == "overall"]
cat(sprintf(
  paste0(
    "**On the stations the 2022 CNEFE street/neighborhood tables currently win** ",
    "(%s stations, of which geocodebr resolves %s): median error **%.3f km** ",
    "(geocodebr) vs **%.3f km** (pipeline), a difference of **%+.3f km**. ",
    "Within 500 m: **%.1f%%** vs **%.1f%%**, a difference of **%+.1f pp**."
  ),
  format(gate$n_stations, big.mark = ","), format(gate$n_geocodebr, big.mark = ","),
  gate$median_km_geocodebr, gate$median_km_model, gate$delta_median_km,
  gate$within_500m_geocodebr, gate$within_500m_model, gate$delta_within_500m
))
```

```{r}
#| label: geocodebr-headtohead
gb_cmp[stratum != "geocodebr_tier"] |>
  head_to_head_gt(
    title = "geocodebr vs the pipeline's selected coordinate",
    subtitle = paste(
      "all_covered = every TSE-covered station;",
      "cnefe22_winner = those the 2022 CNEFE street/neighborhood tables currently win."
    )
  ) |>
  cols_label(
    stratum = "Cut", n_stations = "N", n_geocodebr = "N geocodebr", n_model = "N model",
    n_both = "N both", suppressed = "Below N floor"
  ) |>
  tab_source_note(
    "Δ is geocodebr minus pipeline: negative median and positive <500 m favour geocodebr."
  )
```

### By `geocodebr` precision tier

`geocodebr` labels each result with the rung of its cascade it reached, from an exact house
number down to the municipal centroid. The tier decides how much of the substitution is
actually available: if `geocodebr` only reaches parity at street level or better, it cannot
replace the tables on the stations that fall to coarser rungs — which is precisely where the
bespoke aggregates are doing work. `sem_resultado` is the stations `geocodebr` returned
nothing for.

```{r}
#| label: geocodebr-tier
gb_cmp[stratum == "geocodebr_tier", .(
  universe, level, n_stations, n_both,
  median_km_geocodebr, median_km_model, delta_median_km,
  within_500m_geocodebr, within_500m_model, delta_within_500m
)] |>
  head_to_head_gt(title = "Head-to-head by geocodebr precision tier") |>
  cols_label(n_stations = "N", n_both = "N both") |>
  tab_stubhead(label = "Tier")
```

```{r}
#| label: geocodebr-standalone
gb_acc[stratum %in% c("overall", "urban_rural", "region", "match_source")] |>
  acc_ladder_gt(
    title = "geocodebr on its own, over the whole covered set",
    subtitle = "Same metric ladder as the pipeline table above. `match_source` is the precision tier."
  )
```

## TSE coverage density (by year × state)

Each accuracy number should be read against how much ground truth its stratum has.
TSE coordinates begin with the 2018 vintage; pre-2018 station-years have no TSE
reference and are unmeasurable.

```{r}
#| label: coverage
cov_wide <- dcast(cov, sg_uf ~ ano, value.var = "coverage_pct")
cov_wide |>
  gt(rowname_col = "sg_uf") |>
  fmt_number(columns = where(is.numeric), decimals = 1) |>
  sub_missing(missing_text = "—") |>
  tab_header(title = "TSE coverage (% of covered stations)") |>
  tab_source_note("Share of polling stations with a non-missing TSE coordinate.")
```

## `pred_dist` calibration

### Rank-and-filter

Sorting covered stations by predicted error and dropping the worst-predicted tail
should monotonically lower realized median error and raise %-within-500 m. This
demonstrates the ranking carries information even where it is not calibrated in
absolute metres.

```{r}
#| label: rank-filter
rf <- as.data.table(calibration_check$rank_filter)
rf |>
  gt() |>
  fmt_number(columns = c("retained_frac"), decimals = 2) |>
  fmt_number(columns = c("median_km"), decimals = 3) |>
  fmt_number(columns = c("within_500m"), decimals = 1) |>
  fmt_integer(columns = "n_retained") |>
  cols_label(
    drop_frac = "Worst-predicted dropped", retained_frac = "Retained frac",
    n_retained = "N retained", median_km = "Median km", within_500m = "<500m %"
  ) |>
  tab_header(title = "Rank-and-filter on predicted error")

ggplot(rf, aes(x = drop_frac)) +
  geom_line(aes(y = median_km), color = "#007bff") +
  geom_point(aes(y = median_km), color = "#007bff") +
  labs(x = "Fraction of worst-predicted stations dropped",
       y = "Realized median error (km)",
       title = "Dropping the worst-predicted tail lowers realized error")
```

### Reliability & ENCE

```{r}
#| label: reliability
rel <- as.data.table(calibration_check$reliability)
ence <- calibration_check$ence
```

Expected Normalized Calibration Error (ENCE): **`r sprintf("%.3f", ence)`**
(0 = perfectly calibrated point estimate).

```{r}
#| label: reliability-plot
lim <- range(c(rel$mean_pred, rel$mean_realized), na.rm = TRUE)
ggplot(rel, aes(x = mean_pred, y = mean_realized)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(aes(size = n), color = "#007bff") +
  coord_equal(xlim = lim, ylim = lim) +
  labs(x = "Mean predicted error (km)", y = "Mean realized error (km)",
       size = "N", title = "Reliability: predicted vs realized error by bin")
```

## Google reference-validation (spec section 8)

Quantifies Google's own error budget relative to field-GPS TSE, on a covered-only
sample, *before* trusting Google for the uncovered subset. This is a
reference-validation, **not** a release gate and **not** a pipeline accuracy
number. The Google results are a frozen artifact (`data/google_geocoded.csv`);
regenerate it once with `scripts/generate_google_reference.R` (needs a Google
Geocoding API key).

```{r}
#| label: google
#| output: asis
google_file <- "../data/google_geocoded.csv"
if (!file.exists(google_file)) google_file <- "data/google_geocoded.csv"

if (file.exists(google_file)) {
  g <- fread(google_file)
  cat(sprintf(
    "Frozen Google artifact found: **%s** points. See table below.\n\n",
    format(nrow(g), big.mark = ",")
  ))
} else {
  cat(paste0(
    "_No frozen Google artifact found. This section is skipped until ",
    "`scripts/generate_google_reference.R` is run once to produce ",
    "`data/google_geocoded.csv`._\n"
  ))
}
```

```{r}
#| label: google-table
if (file.exists(google_file) && "google_long" %in% names(g) &&
    all(c("tse_long", "tse_lat") %in% names(g))) {
  g[, google_dist := geosphere::distHaversine(
    cbind(tse_long, tse_lat), cbind(google_long, google_lat)
  ) / 1000]
  by_col <- if ("urban_rural" %in% names(g)) "urban_rural" else NULL
  g_summary <- g[, .(
    n = .N,
    median_km = median(google_dist, na.rm = TRUE),
    within_500m = 100 * mean(google_dist <= 0.5, na.rm = TRUE)
  ), by = by_col]
  gt(g_summary) |>
    fmt_number(columns = "median_km", decimals = 3) |>
    fmt_number(columns = "within_500m", decimals = 1) |>
    tab_header(title = "Google-vs-TSE agreement (reference validation)")
}
```
