---
title: "Geocoding Brazilian Polling Stations with Administrative Data Sets"
author: "F. Daniel Hidalgo"
date: "2021-01-22"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-fold: true
    theme: cosmo
    embed-resources: true
    fig-width: 10
    fig-height: 6
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

# Load required libraries
library(targets)
library(data.table)
library(dplyr)
library(gt)
library(sf)
sf::sf_use_s2(FALSE)

```

This document outlines an approach to geocoding Brazilian polling station that heavily relies on administrative datasets. In addition to detailing our approach, we also provide an honest, held-out estimate of the error of our method, measured out-of-fold on the subset of stations with independent TSE coordinates.

Our general approach is to generate a series of potential coordinates from a variety of administrative datasets.
We use a machine learning model trained on a subset of the data with coordinates provided by Supreme Electoral Tribunal (*TSE*) to choose among the candidate coordinates.
Inputs to this model are mostly measures of the quality of string matches between the polling station address and administrative data sources, as well as other characteristics of the address and municipality of the polling station.
For each polling station, we select the coordinates with the predicted smallest error among the possible coordinates.

## Data Sources

To geocode the polling stations, we leverage four main data sources:

-   *Cadastro Nacional de Endereços para Fins Estatísticos* (CNEFE) from the 2010 and the 2022 editions of the Census.
-   *Cadastro Nacional de Endereços para Fins Estatísticos* from the 2017 Agricultural Census.
-   *Catálogo de Escolas* from INEP.
-   The [`geocodebr`](https://ipea.github.io/geocodebr/) geocoder, which resolves an address against IBGE's own address surface.

The CNEFE datasets are national databases of addresses prepared by IBGE for the census and include detailed data on streets and addresses.
The 2010 and 2017 versions includes private addresses, as well as listings of government buildings (such as schools) and the names of local establishments (such as the names of schools or businesses).
The 2017[^1] version only includes agricultural properties.
Addresses in *rural* census tracts (*setores censitários*) in the 2010 CNEFE have longitude and latitude, while all agricultural properties in the 2017 CNEFE are geocoded. The 2022 CNEFE geocodes all addresses.

[^1]: Details on the 2017 CNEFE can be found at this [link](https://biblioteca.ibge.gov.br/visualizacao/livros/liv101638_notas_tecnicas.pdf)

The 2010 Census data did not include coordinates for addresses in urban census tracts.
To partially overcome this issue, we compute the centroid of the census tract and assign this coordinate to each property in the urban census tract.
Because urban census tracts tend to be compact, tract centroid should still be fairly close to the true coordinates.
Nevertheless, this imputation step will lead to more error for urban addresses than rural addresses when the chosen coordinate is from the 2010 CNEFE data.

The INEP data is a catalog of private and public schools with addresses and longitude and latitude.[^2]

[^2]: The data can be found at this [link](https://inepdata.inep.gov.br/analytics/saw.dll?dashboard)

Unlike the three datasets above, `geocodebr` is not matched by string distance. It takes an address as separate fields --- state, municipality, street, house number, neighborhood, and postal code (CEP) --- and resolves it against IBGE's address surface, returning a coordinate directly. It works down a cascade: it uses the most precise combination of fields it can satisfy, and where it cannot satisfy one it falls back to a coarser rung, ending at the municipal centroid. We therefore supply every field the TSE record carries. The house number has to be recovered from the free-text address line, which is not a simple matter of removing digits: Brazilian street names routinely contain numbers ("BR 364", "Avenida 12 de Outubro", "Rua 15"), and rural addresses are often located by highway milepost rather than by house number. The CEP is supplied for the 97% of station-years that carry a usable one, which is often enough to place a station in the right neighborhood even when its street cannot be resolved.

Each result carries the precision tier it was resolved at and an uncertainty radius in meters (`desvio_metros`), which is the geocoder's own statement of how far the returned point may be from the address. The radius spans four orders of magnitude across tiers --- a few meters at house-number precision against roughly 17 km at the municipal centroid --- and it enters the selection model as its own covariate, telling the model how much to trust this source for a given station.

## String Matching

To geocode polling stations, we use fuzzy string matching to match polling stations to coordinates in the administrative datasets by name, address, street, or neighborhood.
This string matching procedure generates several candidate coordinates.
To choose among these possible coordinates, we use a Random Forest model trained on a sample of polling stations with coordinates provided by the election authorities.

The general approach is a follows:

1.  Normalize[^3] name and address of polling station.

2.  Normalize addresses and school names in administrative datasets.

3.  Find the "medioid" (i.e. the median point) for all unique streets and neighborhoods in the CNEFE datasets.

4.  Compute the length-normalized Jaro-Winkler string distance between polling station name and the names of schools in the INEP and  CNEFE data in the same municipality as the polling station.

5.  Compute the string distance between the address of polling stations and address of schools in INPE and  CNEFE data.

6.  Compute the string distance between the street name and neighborhood name of the polling station and street and neighborhood names from the CNEFE datasets.

7.  For each candidate produced by steps 4--6, compute a *further* string distance for every address component the matched reference record happens to carry, not only the component the match was made on: school name, street, neighborhood, and (where the source stores a single unsplit address line) the whole address. A CNEFE school record, for example, is a single establishment and so carries a street and a neighborhood alongside its name; scoring all three reveals a school whose name matches perfectly but which sits in the wrong neighborhood. Where a source has no such field --- the street and neighborhood reference tables are coordinate medians over a whole street or neighborhood, and know only the one field they are keyed on --- the corresponding distance is missing, which the boosted tree handles natively.

[^3]: We remove common, but uninformative words, such as "povoado" and "localidade". We standardize common street abbreviations such as replacing "Av" with "Avenida". Finally, for polling station names, we remove words most common in school names, such as "unidade escolar" and "colegio estadual". These are very common, yet not used consistently and as a result, are relatively uninformative. We found that removing them improves matching performance.

The string matching procedure above generates 12 different potential matches.

### Choosing Among Potential Matches

After string matching, we use a boosted tree model to predict the distance between the possible coordinates and the true coordinates.
We treat the coordinates provided by the election authorities as the "ground truth".
This distance is modeled as a function of the following set of covariates:

-   Length-normalized Jaro-Winkler string distance of the match that produced the candidate (string-matched sources only).
-   `geocodebr`'s uncertainty radius, as a separate covariate. It is a distance in kilometers rather than a normalized string distance, so the two do not share a scale and cannot share a column: a six-metre radius would otherwise sort below every possible string distance, and a coarse one above every possible string distance. Each is missing for the candidates the other describes.
-   Four per-component string distances between the polling station and the matched reference record: school name, street, neighborhood, and whole address line (step 7 above). Each is missing when the source carries no such field.
-   Coordinate data source
-   Indicator for whether the address mentions the city center ("centro")
-   Indicator for whether the address mentions being in the countryside (includes the word "rural")
-   Indicator for whether the address mentions a school
-   Indicator for whether the address gives no house number ("s/n"), which is true of roughly a quarter of polling stations and signals informal or rural addressing
-   Log of municipal population
-   Proportion of the population classified as rural
-   Area of the municipality

We use the implementation of the boosted tree model provided in the [`lightgbm`](https://github.com/Microsoft/LightGBM) package and use the [`tidymodels`](https://www.tidymodels.org/) framework for preprocessing and training. We train our model on half of the polling stations with ground truth coordinates and use the other half for testing. We tune the hyperpareters of the model using adaptive resampling[^4] and 10 fold cross-validation.
After tuning, we train the model on all polling stations with ground truth coordinates.
We then use this model to predict the distance between the true coordinates and the candidate coordinates.
For each polling station, we choose the candidate coordinate with the smallest predicted distance.

[^4]: See the `finetune` package reference [materials](https://finetune.tidymodels.org/reference/tune_race_anova.html) for more information on adaptive resampling.

### Example of String Matching

```{r}
#| label: load-data
# When run via tar_render(), targets should be available without store argument
# When run manually, we need to specify the store location

tar_load(
    c(
      inep_string_match, cnefe10_stbairro_match, cnefe22_stbairro_match,
      schools_cnefe10_match, schools_cnefe22_match,
      agrocnefe_stbairro_match, locais, model_predictions,
      tsegeocoded_locais, panel_ids
    ))


best_prediction <- model_predictions |>
  filter(!is.na(dist)) |>
  group_by(local_id) |>
  arrange(local_id, pred_dist, dist) |>
  slice(1)

geocoded_locais <- inner_join(locais, best_prediction) |>
  rename("pred_long" = "long", "pred_lat" = "lat") |>
  left_join(select(tsegeocoded_locais, local_id, tse_lat, tse_long)) |>
  ## if we have ground truth distance from TSE, then assign ground truth
  mutate(
    long = ifelse(is.na(tse_long), pred_long, tse_long),
    lat = ifelse(is.na(tse_lat), pred_lat, tse_lat),
    pred_dist = ifelse(is.na(tse_lat), pred_dist, 0)
  ) |>
  left_join(select(panel_ids, panel_id, local_id)) |>
  mutate(panel_id = ifelse(is.na(panel_id), local_id, panel_id)) |>
  arrange(panel_id, pred_dist) |>
  group_by(panel_id) |>
  mutate(
    panel_lat = first(lat),
    panel_long = first(long)
  ) |>
  relocate(local_id, panel_id, ano, sg_uf, cd_localidade_tse, cod_localidade_ibge,
    .before = everything()
  )
```

To illustrate the string matching procedure, the table below shows the string matching procedure for one polling station where the coordinates are known.
"String distance" is the length-normalized Jaro-Winkler string distance between the address component and its potential match.
"Predicted distance" is the distance from the truth predicted by the gradient-boosted tree model.
The blue row shows the selected match, which is the potential match with the smallest predicted distance.
The last column labeled "Error (km)" is the difference between the known geocoded coordinates and the coordinates from the selected match.

```{r}
#| label: string-match-example
# One randomly chosen TSE-covered station, illustrating how each administrative
# data source proposes a candidate coordinate and how the model ranks them. The
# seed keeps the rendered example stable across builds.
set.seed(1)
example_id <- sample(geocoded_locais$local_id[!is.na(geocoded_locais$tse_lat)], 1)
example_ps <- as.data.frame(geocoded_locais[geocoded_locais$local_id == example_id, ])

# Per-source match tables, filtered to the example station (0 or 1 row each).
frames <- list(
  inep = as.data.frame(inep_string_match[inep_string_match$local_id == example_id, ]),
  cnefe10_sch = as.data.frame(schools_cnefe10_match[schools_cnefe10_match$local_id == example_id, ]),
  cnefe22_sch = as.data.frame(schools_cnefe22_match[schools_cnefe22_match$local_id == example_id, ]),
  cnefe10 = as.data.frame(cnefe10_stbairro_match[cnefe10_stbairro_match$local_id == example_id, ]),
  cnefe22 = as.data.frame(cnefe22_stbairro_match[cnefe22_stbairro_match$local_id == example_id, ]),
  agrocnefe = as.data.frame(agrocnefe_stbairro_match[agrocnefe_stbairro_match$local_id == example_id, ])
)

# Each model match_type maps to the reference dataset it came from: the normalized
# polling-station field that was matched, and where the matched string lives.
# Fail loud if a match table's columns drift rather than silently dropping a row.
match_spec <- tibble::tribble(
  ~match_type, ~Data, ~query_field, ~frame, ~match_col,
  "schools_inep_name", "INEP School Name", "normalized_name", "inep", "match_inep_name",
  "schools_inep_addr", "INEP School Address", "normalized_addr", "inep", "match_inep_addr",
  "schools_cnefe_name_2010", "2010 CNEFE School", "normalized_name", "cnefe10_sch", "match_schools_cnefe",
  "schools_cnefe_name_2022", "2022 CNEFE School", "normalized_name", "cnefe22_sch", "match_schools_cnefe",
  "st_cnefe_2010", "2010 CNEFE Street", "normalized_st", "cnefe10", "match_cnefe_st",
  "st_cnefe_2022", "2022 CNEFE Street", "normalized_st", "cnefe22", "match_cnefe_st",
  "bairro_cnefe_2010", "2010 CNEFE Neighborhood", "normalized_bairro", "cnefe10", "match_cnefe_bairro",
  "bairro_cnefe_2022", "2022 CNEFE Neighborhood", "normalized_bairro", "cnefe22", "match_cnefe_bairro",
  "bairro_cnefe_2017", "2017 CNEFE Neighborhood", "normalized_bairro", "agrocnefe", "match_agrocnefe_bairro",
  "geocodebr", "geocodebr", NA, NA, NA
)

pick_string <- function(frame_key, col) {
  if (is.na(frame_key)) {
    return(NA_character_) # geocodebr returns coordinates directly, no fuzzy string
  }
  df <- frames[[frame_key]]
  if (is.null(df)) stop("Unknown frame key in match_spec: ", frame_key)
  if (!col %in% names(df)) {
    stop(
      "Column '", col, "' missing from '", frame_key,
      "' match table -- schema drifted; update match_spec."
    )
  }
  if (nrow(df) == 0) NA_character_ else as.character(df[[col]][1])
}

# model_predictions is the authoritative per-station candidate set: one row per
# match_type, already carrying the string distance (mindist), coordinates, the
# model's predicted error (pred_dist) and, for TSE-covered stations, the true
# error (dist). Everything but the matched string text comes from here.
ex_pred <- as.data.table(model_predictions)[local_id == example_id]
stopifnot(all(ex_pred$match_type %in% match_spec$match_type))

table_data <- ex_pred[, .(
  match_type,
  `String Distance` = mindist,
  `Predicted Error (km)` = pred_dist,
  `True Error (km)` = dist
)] |>
  inner_join(match_spec, by = "match_type") |>
  rowwise() |>
  mutate(
    # Non-string sources (geocodebr) contribute a coordinate but no string comparison.
    `Polling Station String` = if (is.na(query_field)) NA_character_ else example_ps[[query_field]],
    Match = pick_string(frame, match_col),
    `String Distance` = if (is.na(query_field)) NA_real_ else `String Distance`
  ) |>
  ungroup() |>
  arrange(`Predicted Error (km)`) |>
  mutate(rank = row_number()) |>
  select(
    Data, `Polling Station String`, Match, `String Distance`,
    `Predicted Error (km)`, `True Error (km)`, rank
  )


table_data |>
  gt() |>
  cols_hide(c(rank)) |>
  fmt_number(matches("String Distance|Error"),
    decimals = 2
  ) |>
  tab_header(
    title = "Example of String Matching",
    subtitle = paste0(
      "Polling Station Name is ", example_ps$nm_locvot,
      ". Polling Station Address is ", example_ps$ds_endereco
    )
  ) |>
  tab_style(
    style = cell_fill(color = "lightcyan"),
    locations = cells_body(rows = (rank == 1))
  ) |>
  tab_source_note("Highlighted row is the selected match (lowest predicted error).")
```

## Estimating Geocoding Error

Estimating accuracy honestly is constrained by one fact: wherever the TSE has a
coordinate, our published output *is* that TSE coordinate, and the TSE coordinate
is also the model's training target. The model-selected coordinate reaches the
output only for stations the TSE did **not** cover. So accuracy can only be
measured on the TSE-**covered** subset, and it must be measured out-of-fold — each
covered station scored by a model refit on other municipalities, never on itself —
or the numbers flatter themselves. **The numbers below are therefore an
extrapolation to the uncovered stations the model actually determines; that
uncovered subset is unmeasured, not silently folded into the headline.** These
figures are produced by the evaluation harness targets (`accuracy_tables`,
`calibration_check`, `tse_coverage`; see `R/evaluation.R` and
`docs/specs/2026-07-evaluation-spec.md`) and are rebuilt with the pipeline so this
document stays in lockstep with them. (One caveat: the per-fold refits are
leak-free, but the model hyperparameters were tuned once on the production split
and reused across folds rather than re-tuned per fold, leaving a weak residual
hyperparameter-selection channel — an accepted simplification over nested tuning,
so these figures may be marginally optimistic.)

```{r}
#| label: load-eval-targets
# Read the pipeline-produced evaluation targets (store fallback for manual runs).
tryCatch({
  tar_load(c(accuracy_tables, calibration_check, tse_coverage, baseline_comparison))
}, error = function(e) {
  project_root <- normalizePath(file.path(getwd(), ".."), mustWork = FALSE)
  tar_load(c(accuracy_tables, calibration_check, tse_coverage, baseline_comparison),
           store = file.path(project_root, "_targets"))
})
acc <- as.data.table(accuracy_tables)
overall <- acc[stratum == "overall"]
```

```{r}
#| label: error-intro
#| output: asis
cat(sprintf(
  paste0(
    "On the covered set, the out-of-fold median error is **%.2f km** and ",
    "**%.1f%%** of stations land within 500 m of their TSE coordinate, at a ",
    "match rate of **%.1f%%** over %s covered stations. Positional error is a ",
    "right-skewed distribution, so we report percentiles and within-threshold ",
    "shares rather than a single mean, and always alongside the match rate ",
    "(the two trade off)."
  ),
  overall$median_km, overall$within_500m, overall$match_rate,
  format(overall$n_total, big.mark = ",")
))
```

The table below cuts accuracy by urban/rural census tract and by region; rural
stations are multiples worse than urban, the dominant axis of variation. Strata
below the held-out sample-size floor are suppressed (counts shown, medians not).

```{r}
#| label: error-table
acc[stratum %in% c("overall", "urban_rural", "region")][, .(
  Stratum = stratum, Level = level, `N covered` = n_total,
  `Match %` = match_rate, `Median km` = median_km,
  `<500m %` = within_500m, `<1km %` = within_1km
)] |>
  gt(groupname_col = "Stratum", rowname_col = "Level") |>
  fmt_number(columns = c("Median km"), decimals = 2) |>
  fmt_number(columns = c("Match %", "<500m %", "<1km %"), decimals = 1) |>
  fmt_integer(columns = "N covered") |>
  sub_missing(missing_text = "—") |>
  tab_header(title = "Out-of-fold geocoding error (TSE-covered set)") |>
  tab_source_note("Error in kilometres. Suppressed strata show counts only.")
```

```{r}
#| label: calibration-summary
#| output: asis
rf <- as.data.table(calibration_check$rank_filter)
cat(sprintf(
  paste0(
    "The pipeline's own `pred_dist` score is a usable ranking of error even ",
    "where it is not calibrated in absolute metres: dropping the worst-predicted ",
    "%.0f%% of stations lowers the realized median error from %.2f km to %.2f km ",
    "and raises the share within 500 m from %.1f%% to %.1f%%. The absolute ",
    "calibration gap (Expected Normalized Calibration Error) is %.3f. ",
    "Prediction-interval coverage is deferred to the calibrated-quantile upgrade."
  ),
  100 * max(rf$drop_frac),
  rf[drop_frac == 0, median_km], rf[drop_frac == max(drop_frac), median_km],
  rf[drop_frac == 0, within_500m], rf[drop_frac == max(drop_frac), within_500m],
  calibration_check$ence
))
```

### Is the selection model doing the work?

A model that barely beats a one-line rule is not paying for its complexity. We
score a trivial deterministic alternative through the same protocol: for each
station, take the highest-precedence candidate available — INEP school matched on
name, then CNEFE school, then INEP matched on address, then a geocoded address,
then the matched street's median coordinate, then the neighborhood's centroid —
breaking ties within a rank on the closest string match. It trains on nothing.
Both selectors rank the same candidates, so they geocode exactly the same
stations and the comparison is pure accuracy.

```{r}
#| label: baseline-summary
#| output: asis
cmp_overall <- as.data.table(baseline_comparison)[stratum == "overall"]
cat(sprintf(
  paste0(
    "On the covered set, the heuristic reaches a median error of **%.2f km** ",
    "with **%.1f%%** of stations within 500 m, against **%.2f km** and ",
    "**%.1f%%** for the selection model — a difference of **%+.2f km** in ",
    "median error and **%+.1f percentage points** within 500 m (a negative ",
    "and a positive, respectively, favour the model)."
  ),
  cmp_overall$median_km_baseline, cmp_overall$within_500m_baseline,
  cmp_overall$median_km_model, cmp_overall$within_500m_model,
  cmp_overall$delta_median_km, cmp_overall$delta_within_500m
))
```

The full per-stratum comparison is in the evaluation report
(`reports/evaluation_report.qmd`).

A separate, covered-only Google reference-validation (Google-vs-TSE agreement) is
reported in the exploratory evaluation report (`reports/evaluation_report.qmd`)
when its frozen artifact is present; it quantifies Google's error budget and is
not a pipeline accuracy number.
