---
title: "Geocoding Brazilian Polling Stations with Administrative Data Sets"
author: "F. Daniel Hidalgo"
date: "2026-08-14"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-fold: true
    theme: cosmo
    embed-resources: true
    fig-width: 10
    fig-height: 6
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

# Load required libraries
library(targets)
library(data.table)
library(dplyr)
library(gt)
library(sf)
sf::sf_use_s2(FALSE)
```

This document outlines an approach to geocoding Brazilian polling station that heavily relies on administrative datasets. In addition to detailing our approach, we also provide an honest, held-out estimate of the error of our method, measured out-of-fold on the subset of stations with independent TSE coordinates.

Our general approach is to generate a series of potential coordinates from a variety of administrative datasets.
We use a machine learning model trained on a subset of the data with coordinates provided by Supreme Electoral Tribunal (*TSE*) to choose among the candidate coordinates.
Inputs to this model are mostly measures of the quality of string matches between the polling station address and administrative data sources, as well as other characteristics of the address and municipality of the polling station.
For each polling station, we select the coordinates with the predicted smallest error among the possible coordinates.

## Data Sources

To geocode the polling stations, we leverage four main data sources:

-   *Cadastro Nacional de Endereços para Fins Estatísticos* (CNEFE) from the 2010 and the 2022 editions of the Census.
-   *Cadastro Nacional de Endereços para Fins Estatísticos* from the 2017 Agricultural Census.
-   *Catálogo de Escolas* from INEP.
-   The [`geocodebr`](https://ipea.github.io/geocodebr/) geocoder, which resolves an address against IBGE's own address surface.

The CNEFE datasets are national databases of addresses prepared by IBGE for the census and include detailed data on streets and addresses.
The 2010 and 2017 versions includes private addresses, as well as listings of government buildings (such as schools) and the names of local establishments (such as the names of schools or businesses).
The 2017[^1] version only includes agricultural properties.
Addresses in *rural* census tracts (*setores censitários*) in the 2010 CNEFE have longitude and latitude, while all agricultural properties in the 2017 CNEFE are geocoded. The 2022 CNEFE geocodes all addresses.

[^1]: Details on the 2017 CNEFE can be found at this [link](https://biblioteca.ibge.gov.br/visualizacao/livros/liv101638_notas_tecnicas.pdf)

The 2010 Census data did not include coordinates for addresses in urban census tracts.
To partially overcome this issue, we compute the centroid of the census tract and assign this coordinate to each property in the urban census tract.
Because urban census tracts tend to be compact, tract centroid should still be fairly close to the true coordinates.
Nevertheless, this imputation step will lead to more error for urban addresses than rural addresses when the chosen coordinate is from the 2010 CNEFE data.

The INEP data is a catalog of private and public schools with addresses and longitude and latitude.[^2]

[^2]: The data can be found at this [link](https://inepdata.inep.gov.br/analytics/saw.dll?dashboard)


## String Matching

To geocode polling stations, we use fuzzy string matching to match polling stations to coordinates in the administrative datasets by name, address, street, or neighborhood.
This string matching procedure generates several candidate coordinates.
To choose among these possible coordinates, we use a boosted tree model trained on a sample of polling stations with coordinates provided by the election authorities.

The general approach is a follows:

1.  Normalize[^3] name and address of polling station.

2.  Normalize addresses and school names in administrative datasets.

3.  Find the "medioid" (i.e. the median point) for all unique streets and neighborhoods in the CNEFE datasets.

4.  Compute the Jaccard string distance over character trigrams between polling station name and the names of schools in the INEP and  CNEFE data in the same municipality as the polling station.

5.  Compute the string distance between the address of polling stations and address of schools in INEP and CNEFE data.

6.  Compute the string distance between the street name and neighborhood name of the polling station and street and neighborhood names from the CNEFE datasets.

7.  For each candidate produced by steps 4--6, compute a *further* string distance for every address component the matched reference record happens to carry, not only the component the match was made on: school name, street, neighborhood, and the whole address.

[^3]: We remove common, but uninformative words, such as "povoado" and "localidade". We standardize common street abbreviations such as replacing "Av" with "Avenida". Finally, for polling station names, we remove words most common in school names, such as "unidade escolar" and "colegio estadual". These are very common, yet not used consistently and as a result, are relatively uninformative. We found that removing them improves matching performance.

The string matching procedure above generates up to ten potential matches; together with the coordinate `geocodebr` resolves directly, each polling station has up to eleven candidate coordinates.

### Choosing Among Potential Matches

After string matching, we use boosted tree models to predict the distance between the possible coordinates and the true coordinates.
We treat the coordinates provided by the election authorities as the "ground truth".
Two questions have to be answered about each candidate, and they are not the same question: which candidate is likely to be closest, and how far off the one we ship might be.
We therefore fit two models on the same features and the same training split, differing only in what they estimate.
The *selection* model estimates a candidate's expected error and is what ranks candidates; the *bound* model estimates the 90th percentile of that error and, after a conformal correction, becomes the published `conf_dist_km`.
Both models are fit on the following set of covariates:

-   Trigram string distance of the match that produced the candidate (string-matched sources only).
-   `geocodebr`'s uncertainty radius, as a separate covariate. It is a distance in kilometers rather than a normalized string distance, so the two do not share a scale and cannot share a column: a six-metre radius would otherwise sort below every possible string distance, and a coarse one above every possible string distance. Each is missing for the candidates the other describes.
-   Four per-component string distances between the polling station and the matched reference record: school name, street, neighborhood, and whole address line (step 7 above). Each is missing when the source carries no such field.
-   Coordinate data source
-   Measures of where the station's *other* candidates fall, which let the model treat independent references landing on the same spot as evidence and a lone outlier as suspect: how many candidates the station has at all; the distance to the nearest candidate drawn from a different reference dataset, and which dataset that is; how many different datasets fall within 500 metres; the spread of the station's whole candidate cloud (the median distance between its pairs); and how far this candidate sits from the cloud's centre. Agreement is counted across the five underlying datasets rather than the individual matches.
-   Indicator for whether the address mentions the city center ("centro")
-   Indicator for whether the address mentions being in the countryside (includes the word "rural")
-   Indicator for whether the address mentions a school
-   Indicator for whether the address gives no house number ("s/n"), which is true of roughly a quarter of polling stations and signals informal or rural addressing
-   Log of municipal population
-   Proportion of the population classified as rural
-   Area of the municipality

We use the implementation of the boosted tree model provided in the [`lightgbm`](https://github.com/Microsoft/LightGBM) package and use the [`tidymodels`](https://www.tidymodels.org/) framework for preprocessing and training. We split the polling stations with ground truth coordinates in half by municipality, training on one half and holding out the other. We tune the hyperparameters of each model using adaptive resampling[^4] and 10 fold cross-validation, each on the loss it is trained for: squared error for the selection model, pinball loss for the bound.
The held-out half then serves as the calibration set for the conformal correction that turns the bound model's prediction into `conf_dist_km`.
That correction is computed on each station's *selected* candidate rather than on all of them, so the published bound covers the coordinate that actually ships.
For each polling station, we choose the candidate coordinate with the smallest expected error under the selection model.

[^4]: See the `finetune` package reference [materials](https://finetune.tidymodels.org/reference/tune_race_anova.html) for more information on adaptive resampling.

### Example of String Matching

```{r}
#| label: load-data
# geocoded_locais is the pipeline's own output, carrying the candidate finalize_coords()
# selected (pred_long/pred_lat) and the TSE coordinate where one exists. Selection is read
# from it rather than recomputed here, so this document cannot drift from what shipped.
tar_load(
  c(
    inep_string_match,
    cnefe10_stbairro_match,
    cnefe22_stbairro_match,
    schools_cnefe10_match,
    schools_cnefe22_match,
    agrocnefe_stbairro_match,
    locais,
    model_predictions,
    geocoded_locais
  )
)
geocoded_locais <- as.data.table(geocoded_locais)
```

To illustrate the string matching procedure, the table below shows the string matching procedure for one polling station where the coordinates are known.
"String distance" is the Jaccard distance over character trigrams between the address component and its potential match.
"Expected error" is the error the selection model predicts for that candidate, and "Error bound" is the calibrated 90% upper bound the bound model produces for it.
The blue row shows the selected match, which is the candidate with the smallest expected error.
Its error bound is the `conf_dist_km` published for the station, and it is not necessarily the smallest bound in the table --- the two models are ranking on different quantities.
The last column labeled "True error (km)" is the distance between the known TSE coordinates and the coordinates from each candidate.

```{r}
#| label: string-match-example
# One randomly chosen TSE-covered station that the model also geocoded, illustrating how
# each administrative data source proposes a candidate coordinate and how the model ranks
# them. The seed keeps the rendered example stable across builds. The normalized address
# fields come from locais, since finalize_coords() drops them from the output.
set.seed(1)
example_id <- sample(
  geocoded_locais[!is.na(tse_lat) & !is.na(pred_lat), local_id],
  1
)
example_ps <- as.data.frame(locais[locais$local_id == example_id, ])
example_pick <- geocoded_locais[local_id == example_id]

# Per-source match tables, filtered to the example station (0 or 1 row each).
frames <- list(
  inep = as.data.frame(inep_string_match[inep_string_match$local_id == example_id, ]),
  cnefe10_sch = as.data.frame(schools_cnefe10_match[schools_cnefe10_match$local_id == example_id, ]),
  cnefe22_sch = as.data.frame(schools_cnefe22_match[schools_cnefe22_match$local_id == example_id, ]),
  cnefe10 = as.data.frame(cnefe10_stbairro_match[cnefe10_stbairro_match$local_id == example_id, ]),
  cnefe22 = as.data.frame(cnefe22_stbairro_match[cnefe22_stbairro_match$local_id == example_id, ]),
  agrocnefe = as.data.frame(agrocnefe_stbairro_match[agrocnefe_stbairro_match$local_id == example_id, ])
)

# Each model match_type maps to the reference dataset it came from: the normalized
# polling-station field that was matched, and where the matched string lives.
# Fail loud if a match table's columns drift rather than silently dropping a row.
match_spec <- tibble::tribble(
  ~match_type               , ~Data                     , ~query_field        , ~frame        , ~match_col            ,
  "schools_inep_name"       , "INEP School Name"        , "normalized_name"   , "inep"        , "match_inep_name"     ,
  "schools_inep_addr"       , "INEP School Address"     , "normalized_addr"   , "inep"        , "match_inep_addr"     ,
  "schools_cnefe_name_2010" , "2010 CNEFE School"       , "normalized_name"   , "cnefe10_sch" , "match_schools_cnefe" ,
  "schools_cnefe_name_2022" , "2022 CNEFE School"       , "normalized_name"   , "cnefe22_sch" , "match_schools_cnefe" ,
  "st_cnefe_2010"           , "2010 CNEFE Street"       , "normalized_st"     , "cnefe10"     , "match_st"            ,
  "st_cnefe_2022"           , "2022 CNEFE Street"       , "normalized_st"     , "cnefe22"     , "match_st"            ,
  "st_agrocnefe_2017"       , "2017 CNEFE Street"       , "normalized_st"     , "agrocnefe"   , "match_st"            ,
  "bairro_cnefe_2010"       , "2010 CNEFE Neighborhood" , "normalized_bairro" , "cnefe10"     , "match_bairro"        ,
  "bairro_cnefe_2022"       , "2022 CNEFE Neighborhood" , "normalized_bairro" , "cnefe22"     , "match_bairro"        ,
  "bairro_agrocnefe_2017"   , "2017 CNEFE Neighborhood" , "normalized_bairro" , "agrocnefe"   , "match_bairro"        ,
  "geocodebr"               , "geocodebr"               , NA                  , NA            , NA
)

pick_string <- function(frame_key, col) {
  if (is.na(frame_key)) {
    return(NA_character_) # geocodebr returns coordinates directly, no fuzzy string
  }
  df <- frames[[frame_key]]
  if (is.null(df)) {
    stop("Unknown frame key in match_spec: ", frame_key)
  }
  if (!col %in% names(df)) {
    stop(
      "Column '",
      col,
      "' missing from '",
      frame_key,
      "' match table -- schema drifted; update match_spec."
    )
  }
  if (nrow(df) == 0) NA_character_ else as.character(df[[col]][1])
}

# model_predictions is the authoritative per-station candidate set: one row per
# match_type, already carrying the string distance (mindist), coordinates, the selection
# model's score (pred_logmean), the calibrated distance bound (conf_dist_km) and, for
# TSE-covered stations, the true error (dist). Everything but the matched string text
# comes from here.
ex_pred <- as.data.table(model_predictions)[local_id == example_id]
stopifnot(all(ex_pred$match_type %in% match_spec$match_type))

# The selection model is fit on log(dist + 1e-4); the offset only keeps log() finite at
# dist == 0 and is four orders of magnitude below the precision shown here.
ex_pred[, expected_km := pmax(exp(pred_logmean) - 1e-4, 0)]

table_data <- ex_pred[, .(
  match_type,
  `String Distance` = mindist,
  `Expected Error (km)` = expected_km,
  `Error Bound (km)` = conf_dist_km,
  `True Error (km)` = dist
)] |>
  inner_join(match_spec, by = "match_type") |>
  rowwise() |>
  mutate(
    # Non-string sources (geocodebr) contribute a coordinate but no string comparison.
    `Polling Station String` = if (is.na(query_field)) NA_character_ else example_ps[[query_field]],
    Match = pick_string(frame, match_col),
    `String Distance` = if (is.na(query_field)) NA_real_ else `String Distance`
  ) |>
  ungroup() |>
  # The pipeline's rule: lowest expected error wins (select_best_candidate(), R/model.R).
  arrange(`Expected Error (km)`) |>
  mutate(rank = row_number()) |>
  select(
    Data,
    `Polling Station String`,
    Match,
    `String Distance`,
    `Expected Error (km)`,
    `Error Bound (km)`,
    `True Error (km)`,
    rank
  )

# Fail loud if the row highlighted below is not the one the pipeline actually shipped.
winner <- ex_pred[order(pred_logmean)][1]
stopifnot(
  "highlighted candidate is not the pipeline's pick" = isTRUE(all.equal(
    c(winner$long, winner$lat),
    c(example_pick$pred_long, example_pick$pred_lat)
  ))
)


table_data |>
  gt() |>
  cols_hide(c(rank)) |>
  fmt_number(matches("String Distance|Error"), decimals = 2) |>
  tab_header(
    title = "Example of String Matching",
    subtitle = paste0(
      "Polling Station Name is ",
      example_ps$nm_locvot,
      ". Polling Station Address is ",
      example_ps$ds_endereco
    )
  ) |>
  tab_style(
    style = cell_fill(color = "lightcyan"),
    locations = cells_body(rows = (rank == 1))
  ) |>
  tab_source_note("Highlighted row is the selected match (lowest expected error).")
```

## Estimating Geocoding Error

Estimating accuracy honestly is constrained by one fact: wherever the TSE has a
coordinate, our published output *is* that TSE coordinate, and the TSE coordinate
is also the model's training target. The model-selected coordinate reaches the
output only for stations the TSE did **not** cover. So accuracy can only be
measured on the TSE-**covered** subset, and it must be measured out-of-fold — each
covered station scored by a model refit on other municipalities.

```{r}
#| label: load-eval-targets
# Read the pipeline-produced evaluation targets (store fallback for manual runs).
tryCatch(
  {
    tar_load(c(accuracy_tables, calibration_check, tse_coverage, baseline_comparison, geocodebr_vs_model))
  },
  error = function(e) {
    project_root <- normalizePath(file.path(getwd(), ".."), mustWork = FALSE)
    tar_load(
      c(accuracy_tables, calibration_check, tse_coverage, baseline_comparison, geocodebr_vs_model),
      store = file.path(project_root, "_targets")
    )
  }
)
acc <- as.data.table(accuracy_tables)
overall <- acc[stratum == "overall"]
```

```{r}
#| label: error-intro
#| output: asis
cat(sprintf(
  paste0(
    "On the covered set, the out-of-fold median error is **%.2f km** and ",
    "**%.1f%%** of stations land within 500 m of their TSE coordinate, at a ",
    "match rate of **%.1f%%** over %s covered stations. Positional error is a ",
    "right-skewed distribution, so we report percentiles and within-threshold ",
    "shares rather than a single mean, and always alongside the match rate ",
    "(the two trade off)."
  ),
  overall$median_km,
  overall$within_500m,
  overall$match_rate,
  format(overall$n_total, big.mark = ",")
))
```

The table below cuts accuracy by urban/rural census tract and by region; rural
stations are multiples worse than urban, the dominant axis of variation. Strata
below the held-out sample-size floor are suppressed (counts shown, medians not).

```{r}
#| label: error-table
acc[stratum %in% c("overall", "urban_rural", "region")][, .(
  Stratum = stratum,
  Level = level,
  `N covered` = n_total,
  `Match %` = match_rate,
  `Median km` = median_km,
  `<500m %` = within_500m,
  `<1km %` = within_1km
)] |>
  gt(groupname_col = "Stratum", rowname_col = "Level") |>
  fmt_number(columns = c("Median km"), decimals = 2) |>
  fmt_number(columns = c("Match %", "<500m %", "<1km %"), decimals = 1) |>
  fmt_integer(columns = "N covered") |>
  sub_missing(missing_text = "—") |>
  tab_header(title = "Out-of-fold geocoding error (TSE-covered set)") |>
  tab_source_note("Error in kilometres. Suppressed strata show counts only.")
```

```{r}
#| label: calibration-summary
#| output: asis
rf <- as.data.table(calibration_check$rank_filter)
cov_overall <- as.data.table(calibration_check$coverage)[stratum == "overall"]
cat(sprintf(
  paste0(
    "The published `conf_dist_km` claims the true location is within it for at least ",
    "%.0f%% of stations. Out of fold it holds for **%.1f%%**, at a median bound of ",
    "**%.2f km** against a median realized error of %.2f km. It also ranks: dropping ",
    "the %.0f%% of stations with the widest bounds lowers the realized median error ",
    "from %.2f km to %.2f km and raises the share within 500 m from %.1f%% to %.1f%%."
  ),
  calibration_check$nominal,
  cov_overall$coverage,
  cov_overall$median_bound_km,
  cov_overall$median_error_km,
  100 * max(rf$drop_frac),
  rf[drop_frac == 0, median_km],
  rf[drop_frac == max(drop_frac), median_km],
  rf[drop_frac == 0, within_500m],
  rf[drop_frac == max(drop_frac), within_500m]
))
```

### Is the selection model doing the work?

A model that barely beats a one-line rule is not paying for its complexity. We
score a trivial deterministic alternative through the same protocol: for each
station, take the highest-precedence candidate available — INEP school matched on
name, then CNEFE school, then INEP matched on address, then a geocoded address,
then the matched street's median coordinate, then the neighborhood's centroid —
breaking ties within a rank on the closest string match. It trains on nothing.
Both selectors rank the same candidates, so they geocode exactly the same
stations and the comparison is pure accuracy.

```{r}
#| label: baseline-summary
#| output: asis
cmp_overall <- as.data.table(baseline_comparison)[stratum == "overall"]
cat(sprintf(
  paste0(
    "On the covered set, the heuristic reaches a median error of **%.2f km** ",
    "with **%.1f%%** of stations within 500 m, against **%.2f km** and ",
    "**%.1f%%** for the selection model — a difference of **%+.2f km** in ",
    "median error and **%+.1f percentage points** within 500 m (a negative ",
    "and a positive, respectively, favour the model)."
  ),
  cmp_overall$median_km_baseline,
  cmp_overall$within_500m_baseline,
  cmp_overall$median_km_model,
  cmp_overall$within_500m_model,
  cmp_overall$delta_median_km,
  cmp_overall$delta_within_500m
))
```

The full per-stratum comparison is available from the `baseline_comparison`
pipeline target.

### How much of this could an off-the-shelf geocoder do?

`geocodebr` is one of the candidate sources, and it is built on IBGE's own
aggregation of the same 2022 census address file that two of our reference tables
--- matched street segments and matched neighborhoods --- are built from. That
overlap raises a fair question: are those hand-built tables earning their place,
or would the off-the-shelf geocoder do as well on its own? We answer it by
scoring `geocodebr`'s coordinates against the TSE ground truth on the stations
where one of those two tables currently supplies the winning match, comparing
only stations both approaches resolve so a difference in accuracy is not a
difference in coverage in disguise.

```{r}
#| label: geocodebr-gate-summary
#| output: asis
gate <- as.data.table(geocodebr_vs_model)[
  universe == "cnefe22_winner" & stratum == "overall"
]
cat(sprintf(
  paste0(
    "On the %s covered stations the 2022 street and neighborhood tables win, ",
    "`geocodebr` returns a coordinate for %s. Where both resolve the station, ",
    "`geocodebr`'s median error is **%.2f km** against **%.2f km** for the ",
    "pipeline's pick, and **%.1f%%** of its coordinates land within 500 m ",
    "against **%.1f%%** --- a difference of **%+.2f km** and **%+.1f ",
    "percentage points** (a negative and a positive, respectively, favour ",
    "`geocodebr`)."
  ),
  format(gate$n_stations, big.mark = ","),
  format(gate$n_geocodebr, big.mark = ","),
  gate$median_km_geocodebr,
  gate$median_km_model,
  gate$within_500m_geocodebr,
  gate$within_500m_model,
  gate$delta_median_km,
  gate$delta_within_500m
))
```

The breakdown by `geocodebr`'s own precision tier --- how far down its cascade
each result came from --- is available from the `geocodebr_vs_model` pipeline
target, along with the same comparison over the whole covered set.

A separate, covered-only Google reference-validation (Google-vs-TSE agreement)
can be produced by running `scripts/generate_google_reference.R` once to freeze
`data/google_geocoded.csv`; it quantifies Google's error budget and is not a
pipeline accuracy number.
